\chapter{Governing Equations} \label{chap:governing-equations}
\section{Fluid Description for Flow}
In this section, we will derive the governing equations of the flow in magnetic nozzle, starting from the fluid description for plasma.

In magnetic nozzle, the magnetic field is along the nozzle, which we denote as z-axis. Due to Lorentz force, the charged particles gyrates about the magnetic field lines. Because the magnetic moment is invariant in such situation. \cite{chen_introduction_2016} The fluid velocity of particles can be written as $\mathbf{v} = v\mathbf{B}/B$, meaning that the particles move along the magnetic field lines. Therefore the conservation of density 
\[ 
\pdv{n}{t} + \div(n\mathbf{v}) = 0 
\Rightarrow 
\pdv{n}{t} + B\pdv{z}(\frac{nv}{B}) = 0  
\]
In the derivation, the divergence-free condition for magnetic field, $\div{\mathbf{B}} = 0$, is used.

To derive the second governing equation, we start from the conservation of momentum, 
\[ \pdv{v}{t} + v\pdv{v}{z} = -\frac{1}{\rho}\grad{p} \]
Let $\grad{p} = k_BT\pdv*{n}{z}$, we have
\[ \pdv{v}{t} + v\pdv{v}{z} = -c_s^2\frac{1}{n}\pdv{n}{z} \]
where $c_s^2 = k_BT/m$ is the square of sound speed.

Therefore the dynamics of the flow in magnetic nozzle can be characterized by the conservation of density and momentum,
\begin{align*}
	&\pdv{n}{t} + B\pdv{z}(\frac{nv}{B}) = 0\\
	&\pdv{v}{t} + v\pdv{v}{z} = -c_s^2\frac{1}{n}\pdv{n}{z}
\end{align*}
The magnetic field profile was discussed in Sec.\ref{sec:magnetic-field-in-nozzle}.

\section{Linearized Equations}
\subsection{Non-dimensionalization}
For convenience, we nondimensionalize the governing equations by normalizing the velocity to $c_s$, $v\mapsto v/c_s$, $z$ to system length $L$, $z \mapsto z/L$ and time $t\mapsto c_s t/L$. The governing equations become
\begin{align}
    &\pdv{n}{t} + n\pdv{v}{z} + v\pdv{n}{z} - nv\frac{\partial_z B}{B} = 0 
	\label{eq:conservation-of-mass}
	\\
    &n\pdv{v}{t} + nv\pdv{v}{z} = -\pdv{n}{z}
	\label{eq:conservation-of-momentum}
\end{align}
and the nondimensionalized equilibrium condition is
\begin{align}
    &\pdv{z}(\frac{n_0v_0}{B}) = 0 \label{eq:equilibrium-convervation-of-mass}\\
    &v_0\pdv{v_0}{z} = -\frac{1}{n_0}\pdv{n_0}{z} \label{eq:equilibrium-convervation-of-momentum}
\end{align}

\subsection{Linearization}
As illustrated in Sec.\ref{sec:two-stream-instability}, it is essential to linearize the governing equations in order to investigate the instability of plasma. Now we are going to derive the linearized governing equations with the equilibrium conditions given in above.

Let $n = n_0(z) + \tilde{n}(z,t)$ and $v = v_0(z) + \tilde{v}(z,t)$, where $\tilde{n}$ and $\tilde{v}$ are small perturbed quantities.

We first linearize Eq.(\ref{eq:conservation-of-mass}) by setting $n=n_0+\tilde{n}$ and $v=v_0+\tilde{v}$,
\[    \pdv{(n_0+\tilde{n})}{t} 
	+ (n_0+\tilde{n})\pdv{(v_0+\tilde{v})}{z} 
	+ (v_0+\tilde{v})\pdv{(n_0+\tilde{n})}{z} 
	- (n_0+\tilde{n})(v_0+\tilde{v})\frac{\partial_z B}{B} = 0 
\]
By ignoring the second order perturbations, we obtain
\[ \frac{1}{n_0}\pdv{\tilde{n}}{t} 
+ \pdv{v_0}{z} + \frac{\tilde{n}}{n_0}\pdv{v_0}{z} + \pdv{\tilde{v}}{z}
+ \frac{v_0}{n_0}\pdv{n_0}{z} + \frac{\tilde{v}}{n_0}\pdv{n_0}{z} + \frac{v_0}{n_0}\pdv{\tilde{n}}{z} 
- v_0\frac{\partial_z B}{B} - \tilde{v}\frac{\partial_z B}{B} - \tilde{n}\frac{v_0}{n_0}\frac{\partial_z B}{B} = 0 
\]


Using the equilibrium condition Eq.(\ref{eq:equilibrium-convervation-of-mass}), some of the terms are canceled. Moreover, the last term can be written as 
\[ \tilde{n}\frac{v_0}{n_0}\frac{\partial_z B}{B} = \frac{\tilde{n}}{n_0}\left( \frac{\partial_z n_0}{n_0}v_0 + \pdv{v_0}{z} \right) \]
Now, we get the linearized conservation of mass,
\begin{equation} \label{eq:linearized-conservation-of-mass}
	\frac{1}{n_0}\pdv{\tilde{n}}{t} 
        + \pdv{\tilde{v}}{z} + v_0\tilde{Y} + \tilde{v}\frac{\partial_z n_0}{n_0} - \tilde{v}\frac{\partial_z B}{B} = 0 
\end{equation}
where 
\[ \tilde{Y} \equiv \frac{1}{n_0}\pdv{\tilde{n}}{z} - \frac{\partial_z n_0}{n_0^2}\tilde{n} = \pdv{z}(\frac{\tilde{n}}{n_0}) \] 

To linearize the conservation of momentum, we follow the same logic by substituting $n=n_0+\tilde{n}$, and $v=v_0+\tilde{v}$ in Eq.(\ref{eq:conservation-of-momentum}),
\[ (n_0+\tilde{n})\pdv{(v_0+\tilde{v})}{t} + (n_0+\tilde{n})(v_0+\tilde{v})\pdv{(v_0+\tilde{v})}{z} = -\pdv{n}{z} \]

Again, ignore second order perturbations and rearange terms, we have
\[ \pdv{v_0}{t} + v_0\pdv{v_0}{z} + \tilde{v}\pdv{v_0}{z} 
= -\frac{1}{n_0}\pdv{n_0}{z} -\frac{1}{n_0}\pdv{\tilde{n}}{z} -v_0\frac{v_0}{z} - \frac{\tilde{n}}{n_0}v_0\pdv{v_0}{z} \]
Using the equilibrium condition Eq.(\ref{eq:equilibrium-convervation-of-momentum}) on the RHS, we get the linearized conservation of momentum,
\begin{equation} \label{eq:linearized-conservation-of-momentum}
	\pdv{\tilde{v}}{t} + \pdv{(v_0\tilde{v})}{z} = -\tilde{Y}	
\end{equation}

\section{Polynomial Eigenvalue Problem}
We can further simplify the problem by combining Eq.(\ref{eq:linearized-conservation-of-mass}) and Eq.(\ref{eq:linearized-conservation-of-momentum}) into a single equation. We can substitute Eq.(\ref{eq:linearized-conservation-of-momentum}) into Eq.(\ref{eq:linearized-conservation-of-mass}) to eliminate $\tilde{Y}$,

\begin{equation} \label{eq:single-governing-equation}
	\pdv{t}\frac{\tilde{n}}{n_0} 
	+ \pdv{\tilde{v}}{z} - v_0\left(\pdv{t}\tilde{v} 
	+ \pdv{(v_0\tilde{v})}{z}\right) 
	+ \tilde{v}\frac{\partial_z n_0}{n_0} 
	- \tilde{v}\frac{\partial_z B}{B} 
	= 0
\end{equation}

In order to investigate the instability of the flow, we need formulate it as an eigenvalue problem. To do that, we assume the perturbed density and velocity are oscillatory, i.e. $\tilde{n}, \tilde{v} \sim \exp(-i\omega t)$, where $\omega$ is the oscillation frequency of the perturbed quantities. This frequency can be a complex number. 

As illustrated in Sec.\ref{sec:two-stream-instability}, the flow can be stable or unstable depending on the imaginary part of the frequency. If $\Im(\omega) > 0$, then the perturbed quantities $\tilde{n} \sim \exp(\Im(\omega) t)$, which means it grows exponentially with time, hence unstable. If $\Im(\omega) \leq 0$, then the amplitude of the perturbed quanties are either unchanged or exponentially decreasing, hence the flow is stable.

By assuming oscillatory perturbed quantities, Eq.(\ref{eq:single-governing-equation}) becomes,
\begin{equation}
	-i\omega\frac{\tilde{n}}{n_0} 
	+ \pdv{\tilde{v}}{z} - v_0\left(-i\omega\tilde{v} 
	+ \pdv{(v_0\tilde{v})}{z}\right) 
	+ \tilde{v}\frac{\partial_z n_0}{n_0} 
	- \tilde{v}\frac{\partial_z B}{B} 
	= 0
\end{equation}

Using the equilibrium condition Eq.(\ref{eq:equilibrium-convervation-of-mass}), we can eliminate the term $\partial_z B/B$,
\[ 
	-i\omega\frac{\tilde{n}}{n_0} 
	+ \pdv{\tilde{v}}{z} 
	+ v_0\left(i\omega \tilde{v} - v_0\pdv{\tilde{v}}{z} - \tilde{v}\pdv{v_0}{z} \right)
	- \tilde{v}\frac{\partial_z v_0}{v_0} 
	= 0 
\]

Rearrange terms, we have
\[
	-i\omega\frac{\tilde{n}}{n_0} 
	+ i\omega v_0\tilde{v}
	+ (1-v_0^2)\pdv{\tilde{v}}{z} 
	- \left(v_0+\frac{1}{v_0}\right)\pdv{v_0}{z}\tilde{v} = 0	
\]

Now we take $\pdv*{t}$ on Eq.(\ref{eq:linearized-conservation-of-momentum}). Recall the fact that $\tilde{Y} = \partial_z(\tilde{n}/n_0)$, we have
\[ 
	\omega^2\tilde{v} + i\omega\left(v_0\pdv{\tilde{v}}{z} + \tilde{v}\pdv{v_0}{z}\right) 
	= \pdv{t}\pdv{z}(\frac{\tilde{n}}{n_0}) 
\]
Apply $\partial_t$ operator first, we get
\[ 
	\omega^2\tilde{v} + i\omega\left(v_0\pdv{\tilde{v}}{z} + \tilde{v}\pdv{v_0}{fz}\right) 
	= \pdv{z}(-i\omega v_0\tilde{v}
	- (1-v_0^2)\pdv{\tilde{v}}{z} 
	+ \left(v_0+\frac{1}{v_0}\right)\pdv{v_0}{z}\tilde{v})	
\]
Expand the RHS and collect terms, we get
\begin{equation} \label{eq:polynomial-eigenvalue-problem}
	\begin{aligned}
		&\omega^2 \tilde{v} \\ 
		&+2i\omega\left(v_0\pdv{}{z} + \pdv{v_0}{z}\right) \tilde{v} \\ 
		&+\left[ (1-v_0^2)\pdv[2]{}{z} 
		-\left(3v_0 + \frac{1}{v_0}\right)\pdv{v_0}{z}\pdv{}{z} 
		- \left(1-\frac{1}{v_0^2}\right)\left(\pdv{v_0}{z}\right)^2 
		- \left(v_0+\frac{1}{v_0}\right)\pdv[2]{v_0}{z} \right]\tilde{v}
		= 0
	\end{aligned}
\end{equation}

Eq.(\ref{eq:polynomial-eigenvalue-problem}) is an polynomial eigenvalue problem. In Chap.(\ref{chap:methodology}) we will discuss the methods to tackle this problem.