\chapter{Governing Equations}
\section{Governing Equations for Flow in Magnetic Nozzle}
In this section, we will derive the governing equations of the flow in magnetic nozzle, starting from the fluid description for plasma.

In magnetic nozzle, the magnetic field is along the nozzle, which we denote as z-axis. Due to Lorentz force, the charged particles gyrates about the magnetic field lines. Because the magnetic moment is invariant in such situation (\textbf{reference}). The fluid velocity of particles can be written as $\mathbf{v} = v\mathbf{B}/B$, meaning that the particles move along the magnetic field lines. Therefore the conservation of density 
\[ 
\pdv{n}{t} + \div(n\mathbf{v}) = 0 
\Rightarrow 
\pdv{n}{t} + B\pdv{z}(\frac{nv}{B}) = 0  
\]
In the derivation, $\div{\mathbf{B}} = 0$ is used.

To derive the second governing equation, we start from the conservation of momentum, 
\[ \pdv{v}{t} + v\pdv{v}{z} = -\frac{1}{\rho}\grad{p} \]
Let $\grad{p} = k_BT\pdv*{n}{z}$, we have
\[ \pdv{v}{t} + v\pdv{v}{z} = -c_s^2\frac{1}{n}\pdv{n}{z} \]
where $c_s^2 = k_BT/m$ is the square of sound speed.

Therefore the dynamics of the flow in magnetic nozzle can be characterized by the conservation of density and momentum,
\begin{align*}
	&\pdv{n}{t} + B\pdv{z}(\frac{nv}{B}) = 0\\
	&\pdv{v}{t} + v\pdv{v}{z} = -c_s^2\frac{1}{n}\pdv{n}{z}
\end{align*}

\section{Linearized Equations}
For convenience, we nondimensionalize the governing equations by normalizing the velocity to $c_s$, $v\mapsto v/c_s$, $z$ to system length $L$, $z \mapsto z/L$ and time $t\mapsto c_s t/L$. The governing equations become
\begin{align}
    &\pdv{n}{t} + n\pdv{v}{z} + v\pdv{n}{z} - nv\frac{\partial_z B}{B} = 0 \\
    &n\pdv{v}{t} + nv\pdv{v}{z} = -\pdv{n}{z}
\end{align}
and the nondimensionalized equilibrium condition is
\begin{align}
    &\pdv{z}(\frac{n_0v_0}{B}) = 0 \label{eq:equilibrium-convervation-of-mass}\\
    &v_0\pdv{v_0}{z} = -\frac{1}{n_0}\pdv{n_0}{z} \label{eq:equilibrium-convervation-of-momentum}
\end{align}

Now we are going to derive an important intermediate result, the linearized governing equations.
    Let $n = n_0(z) + \tilde{n}(z,t)$ and $v = v_0(z) + \tilde{v}(z,t)$, where $\tilde{n}$ and $\tilde{v}$ are small perturbed quantities. The linearized governing equations are
  \begin{align}
      &\frac{1}{n_0}\pdv{\tilde{n}}{t} 
      + \pdv{\tilde{v}}{z} + v_0\tilde{Y} + \tilde{v}\frac{\partial_z n_0}{n_0} - \tilde{v}\frac{\partial_z B}{B} = 0 
      \label{eq:linearized-conservation-of-mass}
      \\
      &\pdv{\tilde{v}}{t} + \pdv{(v_0\tilde{v})}{z} = -\tilde{Y}
      \label{eq:linearized-conservation-of-momentum}
  \end{align}
  where 
  \[ \tilde{Y} \equiv \frac{1}{n_0}\pdv{\tilde{n}}{z} - \frac{\partial_z n_0}{n_0^2}\tilde{n} = \pdv{z}(\frac{\tilde{n}}{n_0}) \]

\section{Polynomial Eigenvalue Problem}
In order to investigate the instability of magnetic nozzle, we need formulate it as an eigenvalue problem. To do that, we assume the perturbed density and velocity are oscillatory, i.e. $\tilde{n}, \tilde{v} \sim \exp(-i\omega t)$, where $\omega$ is the oscillation frequency of the perturbed quantities. This frequency can be a complex number. If $\omega = \omega_r +i \omega_i$, then the perturbed quantities becomes $\tilde{n} \sim \exp(\omega_i t)\exp(i\omega_r t)$, which means it grows exponentially with time.

Let $\tilde{n}\sim \exp(-i\omega t)$ and $\tilde{v} \sim \exp(-i\omega t)$, then we have the polynomial eigenvalue problem
\begin{equation}
  \begin{aligned}
    &\omega^2 \tilde{v} \\
    +& 2i\omega\left(v_0\pdv{}{z} + \pdv{v_0}{z}\right) \tilde{v} \\
    +& \left[ (1-v_0^2)\pdv[2]{}{z} 
      -\left(3v_0 + \frac{1}{v_0}\right)\pdv{v_0}{z}\pdv{}{z} 
      -\left(1-\frac{1}{v_0^2}\right)\left(\pdv{v_0}{z}\right)^2 
    - \left(v_0+\frac{1}{v_0}\right)\pdv[2]{v_0}{z} \right]\tilde{v}
    = 0
  \end{aligned}
  \label{eq:polynomial-eigenvalue-problem}
\end{equation}


