\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{physics}
\usepackage{float, subcaption, graphicx}

\title{On the Instability of Magnetic Nozzle}
\author{Andrei Smolyakov, Hunt Feng}
\date{\today}

\begin{document}
  \maketitle
  \begin{abstract}
    Linear instability of plasma flow in paraxial magnetic nozzle is studied under the framework of polynomial eigenvalue problem. The linear instability for subsonic and supersonic case are found using spectral method. Spectral pollution is filtered using convergence test. The linear stability for transonic velocity profile is investigaed by employing shooting method. The regular solution to the polynomial eigenvalue problem near the singularity is picked up through Frobenius method. Solutions that are crossing the sonic point smoothly were obtained for transonic case. 
  \end{abstract}

  \section{Introduction}
  \begin{itemize}
    \item Mangetic nozzle.
    \item Magnetic mirror configuration.
    \item Relavent research.
  \end{itemize}
  
  \section{Polynomial Eigenvalue Problem}
  \begin{itemize}
    \item Fluid description.
    \item Equilibrium velocity profile.
    \item Results for subsonic and supersonic case?
  \end{itemize}

  \section{Singular Perturbation}
  \begin{itemize}
    \item Expansion of $\tilde{v}$ near singularity. 
    \item Results for transonic cases.
  \end{itemize}

  \section{Conclusion}
  \begin{itemize}
    \item Breifly talk about the result.
    \item Obtained smooth slutions crossing sonic point.
  \end{itemize}

  \bibliography{references}
  \bibliographystyle{plain}
  \nocite{*}
\end{document}
